---
title: "SQL PRACTICE"
output: html_document
---
*****
## **Filtering Columns and Rows**

With SQL, you can filter columns with SELECT and rows with WHERE. Let's try an example with the Lahman baseball database. First, we need to laod the appropriate packages. 

```{r warning=FALSE,message=FALSE}
library(Lahman)
library(sqldf)
```
*****
Now, suppose we would like to see the homerun totals for the 1927 New York Yankees. We could write the following:

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```
*****
This is how to find all instances where Yankees have hit more than 50 homeruns:
```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA'and HR >= 40 ORDER BY HR DESC"
sqldf(query)
```
*****
This is how to find all players with greater than 40 homeruns but less than 60 strikeouts in a season:
```{r}
query<-"SELECT playerID,yearID,teamID,HR,SO FROM Batting
WHERE HR > 40 and SO < 60"
sqldf(query)
```
*****
This is how to find all Phillies' players who hit more than 30 homeruns in the 1970's:
```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='PHI' and yearID>=1970 and yearID<=1979 and HR > 30"
sqldf(query)
```
*****
## **Order By Syntax**
This portion will explain how to sort results from the queries.

This is how to find every instance of players have more than 50 homeruns, where the players who have the most are located at the top:
```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE HR > 50
ORDER BY HR DESC"
sqldf(query)
```
*****
This is how to find all instances of a player striking out less than 10 times and have at least 400 at-bats. The players with the least strikeouts are at the top:
```{r}
query<-"SELECT playerID,yearID,teamID,HR,SO,AB FROM Batting
WHERE SO<10 and AB>400
ORDER BY SO"
sqldf(query)
```
*****
##**Aggregation**
This portion will explain how to aggregate data in a query. There are many way to aggregate data including:
sum() avg() max() min()

This is how to find Babe Ruth's career homerun totals:
```{r}
query<-"SELECT playerID, sum(HR) FROM Batting
WHERE playerID='ruthba01'
GROUP BY playerID"
sqldf(query)
```
*****
Find the career homerun totals for all players. Limit the display to only those that have a career total of 600 or more. Let's have the players with the highest total at the top:
```{r}
query<-"SELECT playerID, sum(HR) FROM Batting
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"
sqldf(query)
```
*****
Find the players who averaged more than 30 homeruns per season over their career. Put those with the highest averages at the top:
```{r}
query<-"SELECT playerID, avg(HR) FROM Batting
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"
sqldf(query)
```
*****
## **Using Joins to Utilize Multiple Tables**
This portion will illustrate how to join data from multiple data tables into one query.

This is how to create a query with Babe Ruth's full name:
```{r}
query<-"SELECT nameFirst, nameLast, teamID, yearID, HR 
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"
sqldf(query)
```
*****
Find all instances of players hitting more than 50 homeruns. Give first & last names, teamID, yearID, and HR. Let's put players with most HR at the top:
```{r}
query<-"SELECT nameFirst, nameLast, teamID, yearID, HR, weight
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.HR > 50
ORDER BY HR DESC"
sqldf(query)
```
*****
This is how to look at Babe Ruth's career stats with his team name rather than teamID. We are combining three tables into one:
```{r}
query<-"SELECT playerID, name, Batting.yearID, Batting.HR
FROM Batting INNER JOIN Teams
ON Batting.teamID=Teams.teamID AND Batting.yearID=Teams.yearID
WHERE playerID='ruthba01'
ORDER BY Batting.HR DESC"
sqldf(query)
```
*****
This is how to list all the players in the MLB named Bob who averaged more than one million dollars per year in salary for their careers, including last names:
```{r}
query<-"SELECT nameFirst, nameLast, avg(salary)
FROM Salaries INNER JOIN Master
ON Salaries.playerID=Master.playerID
WHERE nameFirst='Bob'
GROUP BY Salaries.playerID
HAVING avg(salary)>1000000
ORDER BY avg(salary) DESC"
sqldf(query)
```

